---
title: "Correlated Numerical Data Analysis"
author: "Azmi"
date: "07/03/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(pacman)
p_load(haven, tidyverse, summarytools, broom)
```



# Tutorial 1

```{r}
myds <- read_dta("5.1.dta")
head(myds)
myds0 <- myds %>% 
  mutate_if(is.labelled, funs(as_factor(.)))
myds0 <- myds0 %>% 
  mutate(gender = factor(female, labels = c("male","female")),
         socialclass = factor(sclass, labels = c("manager_pro", "intermediate", "working", "unclassified")),
         schooltype = factor(schtype, labels = c("independent", "state")),
         urban = factor(schurban, labels = c("urban", "rural")),
         denom = factor(schdenom, labels = c("Catholic", "other"))) %>% 
  select(caseid, schoolid, score, cohort90, gender, socialclass, schooltype, urban, denom)
head(myds0)
```

```{r}
mydsgdenomurban <- myds0 %>% group_by(denom, urban)
ctable(mydsgdenomurban$socialclass, mydsgdenomurban$schooltype)

myds0 %>% group_by(schoolid) %>% summarise(meanscore = mean(score), freq = n()) %>% head()
```

## Null model

```{r}
lmodnull <- lm(score ~ 1, data = myds0)
summary(lmodnull)
```

## Null Model MLE

```{r}
p_load(lme4)
```

```{r}
multlevmodnull <- lmer(score ~ 1 + (1|schoolid), data = myds0, REML = F)
summary(multlevmodnull)
tidy(multlevmodnull)
```

multlevmodnull <- lmer(score ~ 1 + (1|schoolid), data = myds0, REML = F)

something + (1|schooldid) something here is random intercept

(1 + something | schoolid) something here in random slope

at average score is 30.60, when consider the variance of schoolid - with SD of the average 7.81

(compare with mean score, regardless of schoolid, is 31.09)

### compare with null model using loglikelihood

```{r}
logLik(lmodnull)
logLik(multlevmodnull)
```
difference log-likelihood = -2 (LL1 - LL2)

```{r}
-2 * (logLik(multlevmodnull)-logLik(lmodnull))
```


## Student level - cohort (Random Intercept Model)

```{r}
multlevmodcoh_RI <- lmer(score ~ cohort90 + (1|schoolid), data = myds0)
summary(multlevmodcoh_RI)
tidy(multlevmodcoh_RI)
```

cohort90 here can also use as year.
but centering is preferably rather that year

interpret: mean (or average) = 30.56 + 1.21 (cohort90) + random_effect(schoolid)


### Prediction - cohort

```{r}
scoreRI_pred <- augment(multlevmodcoh_RI)
head(scoreRI_pred)
```

.fitted = 30.56 + 1.21(cohort90) + random_effect(schoolid) (see below)

#### random effect

```{r}
ran_multlevmodcoh_RI <- ranef(multlevmodcoh_RI)
head(ran_multlevmodcoh_RI$schoolid)
```

scoreij = b0j + b1cohort90ij + Eij
boj = boo + uoj
boo = 30.56
uoj = random_effect(schoolid)  <- random effect at level 2
    = -6.73  <-- for schoolid 1

score = 30.56 - 6.73 + 1.21 * (-6)


### plot
```{r}
ggplot(scoreRI_pred, aes(cohort90, .fitted, group = schoolid )) + geom_point() + geom_line() + theme_minimal()
```

.fitted <- average score for each schoolid (at each point of cohort)
slope seem parallel <- slope similar for each schoolid


## Student Level - cohort (Random Slope Model)

```{r}
multlevmodcoh_RS <- lmer(score ~ cohort90 + (1 + cohort90 | schoolid), data = myds0, REML = F)
summary(multlevmodcoh_RS)
```


### bobyqa optimize

```{r}
multlevmodcoh_RSb <- lmer(score ~ cohort90 + (1 + cohort90 | schoolid), data = myds0, control = lmerControl(optimizer = "bobyqa"), REML = F)
summary(multlevmodcoh_RSb)
```

### Prediction - with bobyqa optimize

```{r}
scoreRSb_pred <- augment(multlevmodcoh_RSb)
head(scoreRSb_pred)
```

### Compare RI & RS

```{r}
anova(multlevmodcoh_RI, multlevmodcoh_RSb)
```

### plot

```{r}
ra_eff_rs <- ranef(multlevmodcoh_RSb, condVar = T)
head(ra_eff_rs$schoolid)
```

```{r}
ra_eff_rs_scid <- ra_eff_rs$schoolid
ra_eff_rs_scid2 <- dplyr::rename(ra_eff_rs_scid,
                     rs_slope = cohort90,
                     rs_intercept = "(Intercept)")
ggplot(ra_eff_rs_scid2, aes(x = rs_intercept, y = rs_slope)) +
  geom_point() +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  theme_minimal()
```

## Cohort (Random Slope Model) + Gender (assume no random slope)

```{r}
multlevmodcoh_RSgen_noRS <- lmer(score ~ cohort90 + gender + (1 + cohort90 | schoolid), data = myds0, REML = F)
summary(multlevmodcoh_RSgen_noRS)
```


### bobyqa optimizer

```{r}
multlevmodcoh_RSgen_noRSb <- lmer(score ~ cohort90 + gender + (1 + cohort90 | schoolid), data = myds0, REML = F, lmerControl(optimizer = "bobyqa"))
summary(multlevmodcoh_RSgen_noRSb)
```

## Cohort (Random Slope Model) + Gender (random slope)

```{r}
multlevmodcoh_RSgen_RS <- lmer(score ~ cohort90 + gender + (1 + cohort90 + gender | schoolid), data = myds0, REML = F)
summary(multlevmodcoh_RSgen_RS)
```


### Bobyqa optimizer

```{r}
multlevmodcoh_RSgen_RSb <- lmer(score ~ cohort90 + gender + (1 + cohort90 + gender | schoolid), data = myds0, REML = F, lmerControl(optimizer = "bobyqa"))
summary(multlevmodcoh_RSgen_RSb)
```

### comparison

```{r}
anova(multlevmodcoh_RSgen_noRSb, multlevmodcoh_RSgen_RSb)
```

# Level 2 Explanatory

social class as level 2 explanatory

```{r}
freq(myds0$socialclass)
```


## Gender (lv1) Class (lv2)

```{r}
mlmcohRSgendernoRS_classb <- lmer(score ~ cohort90 + gender + socialclass +
                                   (1 + cohort90 | schoolid), data = myds0,
                                  REML = F, lmerControl(optimizer = "bobyqa"))
summary(mlmcohRSgendernoRS_classb)
```

```{r}
mlmcohRSgendernoRS_classb1 <- lmer(score ~ cohort90 + gender + factor(socialclass) +
                                   (1 + cohort90 | schoolid), data = myds0,
                                  REML = F, lmerControl(optimizer = "bobyqa"))
summary(mlmcohRSgendernoRS_classb1)
```

to confirm with prof. why need factor

# Interaction

```{r}
fulmlm_int <- lmer(score ~ cohort90 + gender + socialclass + schooltype + urban +
                     cohort90:schooltype + (1 + cohort90 | schoolid), data = myds0,
                   REML = F, lmerControl(optimizer = "bobyqa"))
summary(fulmlm_int)
```

# Assumption

using gender model

```{r}
summary(multlevmodcoh_RSgen_RSb)
```

```{r}
predscore2 <- augment(multlevmodcoh_RSgen_RSb)
head(predscore2)
```

## Plot 1

```{r}
p_load(merTools)
```

```{r}
RE_multlevmodcoh_RSgen_RSb <- REsim(multlevmodcoh_RSgen_RSb)
plotREsim(RE_multlevmodcoh_RSgen_RSb)
```


## Plot 2

```{r}
p_load(lattice)
```

```{r}
ran_multlevmodcoh_RSgen_RSb <- ranef(multlevmodcoh_RSgen_RSb, condVar = T)
dotplot(ran_multlevmodcoh_RSgen_RSb)
```

```{r}
plot(multlevmodcoh_RSgen_RSb)
qqmath(multlevmodcoh_RSgen_RSb)
```

